Yield Mapping

### 2016-02-10 Rmd/yield-predictors.rmd
Deleted prior code for collecting yields from gridded yield datasets and merging with predictors

# Read in soil-climate grid from Di's work
# ID <- fread(full_path(p_crop, "Soil_clim_match_ID_master.csv"))
# ID[, V1 := NULL]
# setkeyv(ID, c("WTH", "SOL", "Zone"))
# ID[, SharePct := as.numeric(SharePct)]
# zone <- sort(unique(ID$Zone))
# for(i in 1:length(zone)) ID[Zone == zone[i], zoneid := i]
# setkeyv(ID, c("WTH", "Zone"))
# # ID[, FID := 1:.N]
# IDr <- subset(unique(ID[, .(x, y, ID, WTH, Zone, zoneid)]))
# setkeyv(IDr, c("WTH", "Zone"))
# # ID[SOL == "WI_ARBW438"]
# idgr <- dt_to_raster(IDr[, .(x, y, ID, zoneid)], CRS(gcs))
# plot(idgr)

@@

# zmetsamps <- extract(idgr, zmetf, cellnumbers = TRUE)
# zmetxy <- data.frame(zmetf)[, c("x", "y")]
zmetxy <- data.table(data.frame(zmetf)[, 1:3])
wthnms <- substr(zmetxy$nms, 1, 4)
ginds <- grep("g", wthnms)
combis <- expand.grid(data.frame(cbind(LETTERS, LETTERS, LETTERS)))[, 3:1]
combis <- paste0(combis[, 1], combis[, 2], combis[, 3])
wthnms[ginds] <- paste0("G", combis[1:length(ginds)])
zmetxy <- cbind(data.table("WTH" = wthnms), zmetxy)

# xFromCell(idgr, zmetsamps[, "cells"]) - zmetxy$x
# yFromCell(idgr, zmetsamps[, "cells"]) - zmetxy$y
# zmetsamps <- cbind(xyFromCell(idgr, zmetsamps[, "cells"]), zmetsamps)
# zmetsamps <- cbind(zmetxy, zmetsamps)  # harmonize with DSSAT zmetfxy

# zmet_dt <- data.table(data.frame(zmetsamps))[, .(x, y)]
# zmet_dt <- data.table(data.frame(zmetsamps))
# # setkeyv(zmet_dt, c("x", "y"))
# setkeyv(zmet_dt, c("ID", "zoneid"))
# 
# # setkeyv(IDr, c("x", "y"))
# setkeyv(IDr, c("ID", "zoneid"))
# IDr_ss <- merge(IDr, zmet_dt)
# # IDr_ss <- IDr[zmet_dt]  # pick up IDs, soils, wth variables
# IDr_ss[, c("x.x", "y.x", "cells") := NULL]
# setnames(IDr_ss, c("x.y", "y.y"), c("x", "y"))
# setkeyv(IDr_ss, c("WTH", "Zone"))

# Collect soils up here also 
# pts <- IDr_ss
# pts <- cbind.data.frame("x" = 27.4738917, "y" = -16.2783358)  # in Zambia
# coordinates(pts) <- ~x+y
# soilid <- extract(soilgrid, pts)  # grid id corresponding to point
# soilid <- extract(soilgrid, zmetf)  # grid id corresponding to point
# IDr_ss[, soilid := soilid]
# profs <- soilshare[CELL5M %in% soilid]  # profiles at that point
# profs[, SharePct := NULL]
# IDr_ss <- merge(IDr_ss, profs, by.x = "soilid", by.y = "CELL5M")
# IDr_ss[, soilid := NULL]
# setnames(IDr_ss, "SoilProfile", "SOL")
# setkeyv(IDr_ss, c("Zone", "WTH", "SOL"))

# IDr_ss[, plot(x, y)]
# points(zmet_dt, pch = 16, cex = 0.5)
# tst <- extract(idgr, zmet_dt)
# all(tst[order(tst[, 1]), 2] == zmetsamps[order(zmetsamps[, 4]), 5])

# alternative for yield data selection points:
# Creatre stratified, representative random sample of 50 pts from ID/zone grid
# ztab <- freq(idgr$zoneid)[-5, ]
# ztabs <- round((ztab[, 2] / sum(ztab[, 2])) * 50)
# set.seed(34)
# gsamps <- do.call(rbind, lapply(1:4, function(i) {
#   gsamp <- sampleRandom(idgr$zoneid == i, size = ztabs[i], cells = TRUE)
#   sampxy <- xyFromCell(idgr, gsamp[, "cell"])
# }))
# plot(idgr$ID)  # unique(values(idgr$ID))
# length(which(!is.na(values(idgr$ID))))
# length(which(!is.na(unique(values(idgr$ID)))))
# points(gsamps, pch = 16, cex = 0.5)
# 
# # merge those points with the main ID grid
# gsamps_dt <- data.table(gsamps)
# setkeyv(gsamps_dt, c("x", "y"))
# setkeyv(IDr, c("x", "y"))
# IDr_ss <- IDr[gsamps_dt]  # pick up IDs, soils, wth variables
# setkeyv(IDr_ss, c("WTH", "SOL", "Zone"))
# IDr_ss[, plot(x, y)]
# points(gsamps, pch = 16, cex = 0.5)

# sgrid <- fread(fp(p_cropmod, "Soil_clim_match_ID.csv"), 
#                stringsAsFactors = FALSE)
# setnames(sgrid, old = c("V1", "X", "Y", "WeatherGrid", "SoilType"), 
#          new = c("ID", "x", "y", "WTH", "SOL"))
# wgs <- "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"
# sr <- dt_to_raster(sgrid[, list(ID, x, y)], CRSobj = wgs)
# setkeyv(sgrid, c("x", "y"))
# sgrid[SOL == "WI_ARBW438"]
```

# Full DSSAT yield grids
```{r, eval = FALSE}
# # crops <- dir(p_crop, pattern = "maize|soybean", full.names = FALSE)
# crops <- dir(p_crop, pattern = "maize", full.names = FALSE)  # only maize now
# fylds <- lapply(crops, function(X) {  # X <- "maize"
#   fnms <- dir(full_path(p_crop, X), pattern = X)
#   fl <- unlist(lapply(c("S", "R"), function(x) {
#     fnms[grep(paste0(paste0(x, "0"), "|", paste0(x, "9")), fnms)]
#   }))  # reorder to read subsistence first
#   # idt <- copy(IDr)
#   # key(idt)
#   idt <- copy(ID)
#   idt <- subset(unique(idt[, .(x, y, WTH, Zone, SharePct)]))
#   setkeyv(idt, c("WTH", "Zone"))
#   for(y in 1:length(fl)) { # y <- 1
#     # setcolorder(idt, c(4, 7:8, 5:6, 1:3))
#     print(paste("Processing", fl[y]))
#     cr <- fread(input = full_path(p_crop, paste0(X, "/", fl[y])))
#     cr[, "YR" := RUNNO - min(RUNNO) + 1, by = .(TNAM, Zones)]  # unique year  
#     setnames(cr, c("SOIL_ID...", "WSTA....", "Zones"), c("SOL", "WTH", "Zone"))
#     cr[, WTH := substr(WTH, 1, 4)]
#     cr[, c("V1", "RUNNO", "TRNO", "TNAM") := NULL]
#     setkeyv(cr, c("WTH", "Zone"))
#     
#     crm <- merge(cr, idt, by = c("WTH", "Zone"))
# 
#     # nms <- names(crm)[c(6:7, 8, 1:2, 4, 9, 5, 3)]
#     
#     # Soil averaged yields
#     solmu <- crm[, round(weighted.mean(HWAH, w = SharePct)), 
#                  by = .(WTH, Zone, YR)]
#     snm <- gsub("*.*_|\\.csv", "", fl[y])
#     
#     # Yield means
#     v <- solmu[, round(mean(V1)), by = .(WTH, Zone)]
#     setnames(v, "V1", paste0("Y", snm))
#     idt <- idt[v]
#     setkeyv(idt, c("WTH", "Zone"))
#     idt
#   }
#   idt
# })
# names(fylds) <- crops
# 
# rdt <- fylds$maize[, c(1:2, grep("YS", names(fylds$maize))), with = FALSE]
# mzs <- dt_to_raster(rdt, raster()@crs)
# rdt <- fylds$maize[, c(1:2, grep("YR", names(fylds$maize))), with = FALSE]
# mzr <- dt_to_raster(rdt, raster()@crs)
# 
# # par(mar = rep(0, 4))
# plot(calc(mzs, max))
# plot(calc(mzr, mean))
# # plot(calc(soyr, mean))
# # plot(zmetf, pch = 16, cex = 0.5, add = TRUE)

@@

# crops <- dir(p_crop, pattern = "maize", full.names = FALSE)  # only maize now
# # Yield data for those points
# mannms <- c("S", "R")
# ylds <- lapply(crops, function(X) {
#   # X <- crops[1]
#   fnms <- dir(full_path(p_crop, X), pattern = X)
#   fl <- unlist(lapply(mannms, function(x) {
#     fnms[grep(paste0(paste0(x, "0"), "|", paste0(x, "9")), fnms)]
#   }))  # reorder to read subsistence first
#   # fl <- fl[1:3]
#   crml <- lapply(1:length(fl), function(y) { # y <- 1
#     idt <- copy(IDr_ss)
#     print(paste("Processing", fl[y]))
#     cr <- fread(input = full_path(p_crop, paste0(X, "/", fl[y])))
#     cr[, "YR" := RUNNO - min(RUNNO) + 1, by = .(TNAM, Zones)]  # unique year  
#     setnames(cr, c("SOIL_ID...", "WSTA....", "Zones"), c("SOL", "WTH", "Zone"))
#     cr[, WTH := substr(WTH, 1, 4)]
#     cr[, c("V1", "RUNNO", "TRNO", "TNAM") := NULL]
#     setkeyv(cr, c("Zone", "WTH", "SOL"))
# 
#     # merge IDs with subset location
#     # crm <- cr[idt]
#     crm <- merge(cr, idt)
#     # crm[, SharePct := NULL]
#     nms <- names(crm)[c(7:8, 6, 1:2, 9, 3, 5, 4)]
#     setcolorder(crm, nms)
#     mid <- 1:nrow(crm)
#     crm <- cbind(mid, crm)
#     keys <- c("mid", "x", "y", "ID", "WTH", "Zone", "zoneid", "SOL", "YR")
#     setkeyv(crm, keys)
#     # setcolorder(crm, c(nms[grep("HWAH", nms, invert = TRUE)], "HWAH"))
#     newnm <- gsub(".*._||\\.csv", "", fl[y])
#     setnames(crm, "HWAH", newnm)
#     crm
#   })
#   yldmerge <- Reduce(mymerge, crml)
#   ynms <- names(yldmerge)
#   # yldred <- data.table(round(do.call(cbind, lapply(mannms, function(x) {
#     # yldmerge[, rowMeans(.SD), .SDcols = grep(paste0(x, "[0-9]"), ynms)]
#   # }))))
#   # calculate 
#   for(i in mannms) {
#     onms <- list(paste0("YLD", i), grep(paste0(i, "[0-9]"), ynms))
#     # yldmerge[, c(onms[[1]]) := round(rowMeans(.SD)), .SDcols = onms[[2]]]
#     # Selecting rowMax here
#     yldmerge[, c(onms[[1]]) := apply(.SD, 1, max), .SDcols = onms[[2]]]
#     yldmerge[, c(onms[[2]]) := NULL]
#   }
#   
#   # Internal check on yield means
#   if(length(crml) == 8) crind <- list("S" = 1:4, "R" = 5:8, "I" = 9:12) 
#   if(length(crml) == 6) crind <- list("S" = 1:3, "R" = 4:6, "I" = 7:9) 
#   for(ii in sample(1:nrow(yldmerge), size = 20)) {
#     for(jj in 1:length(mannms)) {  # jj = 2
#       a <- yldmerge[ii, ]
#       b <- sapply(crind[[jj]], function(x) {
#         crml[[x]][mid == a$mid, ][[ncol(crml[[x]])]]
#       })
#       #if(a[[paste0("YLD", mannms[jj])]] != round(mean(b))) {
#       if(a[[paste0("YLD", mannms[jj])]] != round(max(b))) {
#         print(paste("Yield mismatch at", ii, mannms[jj]))
#       }
#     }
#   }
#   
#   # swap in averaged yields
#   # yldmerge[, mid := NULL]
#   # yldmerge <- cbind(yldmerge, yldred)
#   yldmerge
# })
# names(ylds) <- crops
# 
# soy_ylds <- fread(fp(p_yield, "DSSAT/soy_ylds.csv"))
# # soy_ylds[, year := as.numeric(substr(PDAT, 1, 4))]
# soy_ylds[, YR := 1:31, by = .(XNAME, TRNO)]  # unique year
# soy_ylds[, c("RUNNO", "PROF", "PDAT", "MDAT", "PRCM") := NULL]
# setnames(soy_ylds, "SOIL_ID...", "SOL")
# 
# # calculate mean and max as YLDS and YLDR
# soy_ylds2 <- soy_ylds[,  round(mean(HWAH)), by = .(XNAME, x, y, WTH, SOL, YR)]
# setnames(soy_ylds2, "V1", "YLDS")
# soy_ylds2[, YLDR := soy_ylds[,  round(max(HWAH)), by = .(XNAME, YR)][, V1]]
# soy_ylds2[, XNAME := NULL]
# 
# ylds$maize[, c("ID", "Zone", "zoneid") := NULL]
# soy_ylds2 <- cbind(c("mid" = 1:nrow(soy_ylds2)), soy_ylds2)
# setnames(soy_ylds2, "V1", "mid")
# 
# ylds <- list(ylds$maize, soy_ylds2)
# names(ylds) <- c("maize", "soybean")

@@

## Soil
### Profile data
Note: Using an updated version of `read_dssat_prof.R` from the rCropMod library. It needs to be rebuilt into that library.  Here it is being sourced before runnign the next section. 
```{r, eval = FALSE}
# pts <- IDr_ss[, .(x, y)]
# pts <- cbind.data.frame("x" = 27.4738917, "y" = -16.2783358)  # in Zambia
# coordinates(pts) <- ~x+y
soilid <- extract(soilgrid, zmetf)  # grid id corresponding to point
# soilid <- extract(soilgrid, pts)  # grid id corresponding to point
profs <- soilshare[CELL5M %in% soilid]  # profiles at that point
profdat <- read_fullSOL_prof(solfile = full_path(p_soil, "WI.SOL"), 
                             profiles = profs$SoilProfile)
setnames(profdat, "prof", "SOL")
solvars <- unique(gsub("[[:digit:]]", "", names(profdat)))
lnms <- names(profdat)

@@

# melt predictors
xy <- unique(ylds$maize[, .(x, y)])
# xy[, ind := 1:.N]
# cols <- c("x", "y")
# xy[, (cols) := lapply(.SD, function(x) round(x, 5)), .SDcols = cols]
# coordinates(xy) <- ~x+y

@@

# Checking merges (disabled at present)
# set.seed(234)
# samp <- sample(1:nrow(regdat$maize), size = 20)
# for(ii in samp) {  # ii <- samp[5]
#   a <- regdat$maize[ii, ]; b <- regdat$soybean[ii, ]
#   d <- ID[WTH == a$WTH & zoneid == a$zoneid & SOL == a$SOL]
#   e <- ylds$maize[WTH == b$WTH & zoneid == b$zoneid & SOL == b$SOL[1] & 
#                   YR == b$YR][1, ]
#   f <- ylds$soybean[WTH == b$WTH & zoneid == b$zoneid & SOL == b$SOL[1] & 
#                     YR == b$YR][1, ]
#   # SOIL climate match
#   for(i in list(b, d, e, f)) {
#     # i <- d
#     for(j in c("zoneid", "WTH", "SOL")) {
#       # print(j)
#       if(a[[j]] != i[[j]]) print(paste("mismatch at", j, "row", ii))
#     }
#     for(j in c("x", "y")) {
#       # print(j)
#       if(round(a[[j]], 5) != round(i[[j]], 5)) {  # round because IDr wasn't
#         print(paste("mismatch at", j, "row", ii))
#       }
#     }
#   }
#   for(i in c("YLDS", "YLDR")) { # i <- "YLDS"
#     if(a[[i]] != e[[i]]) print(paste("maize mismatch at", i, "row", ii))
#     if(b[[i]] != f[[i]]) print(paste("maize mismatch at", i, "row", ii))
#   }
# }

# ylds[[1]][ID == 6968 & zoneid == 1]
# regdat <- lapply(regdat, function(i) {
#   setcolorder(i, c(5, 1:4, 6:ncol(i)))
#   setnames(i, "mid", "midold")
#   i <- cbind(data.table("mid" = 1:nrow(i)), i)
#   i <- setkey(i, "mid")
#   i
# })
# 
# all(regdat$maize$SOL == regdat$soybean$SOL)
# all(regdat$maize$x == regdat$soybean$x)
# all(regdat$maize$y == regdat$soybean$y)
# all(regdat$maize$OP == regdat$soybean$OP)
# all(regdat$maize$SLOC == regdat$soybean$SLOC)

>>>>

@@

zmetxy <- data.table(data.frame(zmetf)[, 1:3])
wthnms <- substr(zmetxy$nms, 1, 4)
ginds <- grep("g", wthnms)
combis <- expand.grid(data.frame(cbind(LETTERS, LETTERS, LETTERS)))[, 3:1]
combis <- paste0(combis[, 1], combis[, 2], combis[, 3])
wthnms[ginds] <- paste0("G", combis[1:length(ginds)])
zmetxy <- cbind(data.table("WTH" = wthnms), zmetxy)

@@

@@

## Soil
### Profile data
Note: Using an updated version of `read_dssat_prof.R` from the rCropMod library. It needs to be rebuilt into that library.  Here it is being sourced before runnign the next section. 
```{r, eval = FALSE}
soilid <- extract(soilgrid, zmetf)  # grid id corresponding to point
profs <- soilshare[CELL5M %in% soilid]  # profiles at that point
profdat <- read_fullSOL_prof(solfile = full_path(p_soil, "WI.SOL"), 
                             profiles = profs$SoilProfile)
setnames(profdat, "prof", "SOL")
solvars <- unique(gsub("[[:digit:]]", "", names(profdat)))
lnms <- names(profdat)

@@

# melt predictors
xy <- unique(ylds$maize[, .(x, y)])

@@
